Crystal as a thermodynamic computer

### 2024-09-05 paper/main.tex
update

\subsection{Energy model}\label{a-2d-surface-programmable-material}

Given that simulating a CA with a deep computational layer on the Ising model requires a large number of spins, making large-scale numerical simulations impractical, we opted to introduce a similar but simpler energy model to observe the behavior on a larger scale.

Denote $S$ to be the length of the material along the ``Space'' axis, $D$
to be the length of the material along the ``Time'' axis. 
Let us define a simpler but equivalent energy model on a 2D lattice of size $S \times D$, whose hamiltonian
is given by
\begin{equation}\label{eq:toy-hamiltonian}
E(\vn) = \sum_{t=1}^{D-1}\sum_{i=1}^{S} (n_{i+1, t} \land \neg(n_{i, t} \land \neg n_{i-1, t})) \veebar n_{i,t+1},
\end{equation}

where $n_{i, t} \in \{0, 1\}$ is the state of the cell at row $i$ and column $t$, the boundary is periodic in the space direction, i.e. $n_{0, t} = n_{S, t}$.
Its ground state with energy $0$ also encodes a $110$ CA, where the direction of computation is from the left to right.
\subsection{Numerical result}

\jinguo{show a comparison from right to left}\jinguo{I have already used D as the dimension}
\ym{Great, i will do this}
Here we conducted numerical experiments to examine both the time complexity along the ``Space'' direction and the ``Time'' direction.
We use binary search to determine the speed of the slowly moving heat bath at which
the simulated annealing procedure would produce 
a ground state configuration with $50\%$ probability. Denote this speed 
as $v$, the simulated annealing procedure will then take $\frac{D}{v}$
stages to complete the computation, each stage we try to flip a spin with 
probability define in (\ref{eq:non-equilibrium-SA}). We name $\frac{D}{v}$ 
as Sweep Time.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/MainNumericalResult.png}
    \caption{Numerical Result}
    \label{fig:main-numerical-result}
\end{figure}

[Fig. \ref{fig:main-numerical-result}(a)] shows the relationship between
$\ln(\ln(D))$ and $\ln(\frac{\text{Sweep Time}}{D})$ shoule be linear, which implies 
the time complexity along the ``Time'' direction should scale as $D\ln(D)^{f(\Lambda)}$. 
We then fit the slopes of each line and $\frac{1}{\ln(\Lambda)}$ and find the 
empirical constant $c$ in (\ref{eq:total-annealing-time}) to be $3$. [Fig. \ref{fig:main-numerical-result}(b)] shows 
the the relationship between Sweep Time and $S$ should be linear. Combining these two results, 
we can conclude that the time complexity should scales as $DS\log(D)^{1+\frac{1}{3\ln(\Lambda)}}$, 
which is consistent with our theoretical prediction in (\ref{eq:total-annealing-time}).
% \begin{figure}[h]
%     \centering
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/TimeComplexityalongTaxis.png}
%         \caption{Time v.s. Depth of the material}
%         \label{fig:image1}
%     \end{minipage}
%     \hspace{0.001\columnwidth} % 两张图片之间的水平间距
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/toymodel_time_vs_width.png}
%         \caption{Time v.s. width of the material}
%         \label{fig:image2}
%     \end{minipage}
% \end{figure}

>>>>

Given that simulating a CA with a deep computational layer on the Ising model requires many spins, making large-scale numerical simulations impractical, we opted to introduce a similar but simpler energy model to observe the behavior on a larger scale.
Let us define a simpler but ``equivalent'' energy model on a 2D lattice of size $S \times D$, whose hamiltonian
is given by
\begin{equation}\label{eq:toy-hamiltonian}
E(\vn) = \sum_{t=1}^{D-1}\sum_{i=1}^{S} (n_{i+1, t} \land \neg(n_{i, t} \land \neg n_{i-1, t})) \veebar n_{i,t+1},
\end{equation}
where $n_{i, t} \in \{0, 1\}$ is the state of the cell at row $i$ and column $t$, the lattice is periodic in the space direction, i.e. $n_{0, t} = n_{S, t}$.
%$S$ to be the length of the material along the ``Space'' axis, $D$
%to be the length of the material along the ``Time'' axis. 
%Its ground state with energy $0$ also encodes a $110$ CA, where the direction of computation is from the left to right.

\jinguo{show a comparison from right to left}\jinguo{I have already used D as the dimension}
\ym{Great, i will do this}

Here we conducted numerical experiments to examine how the cooling time scales with (i)
the ``Space'' dimension $S$, (ii) the ``Time'' dimension $D$, (iii) the ``Width'' of the heat zone $W$, and (iv) the temperature gradient $\Lambda$.
We use the number of SA sweeps as the metric to measure the cooling time, which could be determined with the binary search.
We keep changing the heat bath moving speed $v$ (in unit of columns/sweep) until the system can reach a ground state configuration with $\sim 50\%$ probability.
Let us denote this speed as $v_c$, then the cooling time can be denoted as $\frac{D}{v_c}$.

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/MainNumericalResult.png}
    \caption{Numerical Result\jinguo{images should use the pdf format, the font size should be larger. Is it possible to verify that the Ising model is similar to the toy model? What is $T$ in the xlabel?}}
    \label{fig:main-numerical-result}
\end{figure}

\jinguo{you need to explain the results... what can we learn from the results?}
\jinguo{bad variable name: Sweep Time}
\Cref{fig:main-numerical-result} (a) shows the relationship between
$\ln(\ln(D))$ and $\ln(\frac{\text{Sweep Time}}{D})$ shoule be linear, which implies 
the time complexity along the ``Time'' direction should scale as $D\ln(D)^{f(\Lambda)}$. 
We then fit the slopes of each line and $\frac{1}{\ln(\Lambda)}$ and find the 
empirical constant $c$ in (\ref{eq:total-annealing-time}) to be $3$. [Fig. \ref{fig:main-numerical-result}(b)] shows 
the the relationship between Sweep Time and $S$ should be linear. Combining these two results, 
we can conclude that the time complexity should scales as $DS\log(D)^{1+\frac{1}{3\ln(\Lambda)}}$, 
which is consistent with our theoretical prediction in (\ref{eq:total-annealing-time}).
% \begin{figure}[h]
%     \centering
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/TimeComplexityalongTaxis.png}
%         \caption{Time v.s. Depth of the material}
%         \label{fig:image1}
%     \end{minipage}
%     \hspace{0.001\columnwidth} % 两张图片之间的水平间距
%     \begin{minipage}[b]{0.485\columnwidth}
%         \centering
%         \includegraphics[width=\columnwidth,keepaspectratio]{../notes/images/toymodel_time_vs_width.png}
%         \caption{Time v.s. width of the material}
%         \label{fig:image2}
%     \end{minipage}
% \end{figure}
